---
title: "Online Supplementary Materials"
author: "Juan C. Correa"
date: "2026-07-15"
output: html_document
---

# Appendix A. Data preprocessing and AOI-to-AOI network reconstruction

```{r setup, message=FALSE, warning=FALSE}
library(readr)
library(dplyr)
library(igraph)
library(knitr)
library(ggplot2)
library(pheatmap)
library(nnet)
library(pscl)
```

## Load raw data

```{r load-data}
Secuencia <- read_delim(
  "Secuencia.csv",
  delim = ";",
  escape_double = FALSE,
  trim_ws = TRUE
)

AreasInteres <- read_csv("AreasInteres.csv")

Secuencia <- Secuencia %>%
  left_join(
    AreasInteres,
    by = "Area"
  ) %>%
  dplyr::select(
    -Area
  ) %>%
  rename(
    Area = `Nombre area`
  )
```

## AOI-to-AOI network reconstruction

```{r network-function}
build_aoi_network <- function(df){

  df <- df[order(df$Secuencia), ]

  areas <- df$Area

  if(length(areas) < 2){
    return(NULL)
  }

  ####################################################
  # EARLY VISUAL EXPLORATION
  ####################################################

  FirstAOI <- as.character(
    areas[1]
  )

  FirstTransition <- paste0(
    areas[1],
    "_",
    areas[2]
  )

  First3AOIs <- paste(
    head(areas, 3),
    collapse = "_"
  )

  ####################################################
  # AOI-TO-AOI NETWORK
  ####################################################

  edges <- data.frame(
    from = head(areas, -1),
    to   = tail(areas, -1)
  )

  edges_weighted <- edges %>%
    count(
      from,
      to,
      name = "weight"
    )

  g <- graph_from_data_frame(
    edges_weighted,
    directed = TRUE
  )

  E(g)$weight <- edges_weighted$weight

  ####################################################
  # AOI VISITS
  ####################################################

  node_freq <- table(areas)

  V(g)$visits <- as.numeric(
    node_freq[
      match(
        V(g)$name,
        names(node_freq)
      )
    ]
  )

  ####################################################
  # TRANSITION ENTROPY
  ####################################################

  probs <- edges_weighted$weight /
    sum(edges_weighted$weight)

  Entropy <- -sum(
    probs * log2(probs)
  )

  ####################################################
  # OUTPUT
  ####################################################

  return(
    list(
      graph = g,
      edges = edges,
      edges_weighted = edges_weighted,
      node_freq = node_freq,

      FirstAOI = FirstAOI,
      FirstTransition = FirstTransition,
      First3AOIs = First3AOIs,

      Entropy = Entropy
    )
  )
}
```

## Reconstruct all participant-trial networks

```{r reconstruct-networks}
NetworkList <- list()

TrajectoryFeatures <- data.frame()

subjects <- unique(
  Secuencia$Sujeto
)

for(s in subjects){

  data_s <- filter(
    Secuencia,
    Sujeto == s
  )

  ensayos <- unique(
    data_s$Ensayo
  )

  for(e in ensayos){

    data_se <- filter(
      data_s,
      Ensayo == e
    )

    net <- build_aoi_network(
      data_se
    )

    if(!is.null(net)){

      nombre <- paste0(
        "S",
        s,
        "_E",
        e
      )

      NetworkList[[nombre]] <- net

      TrajectoryFeatures <- rbind(
        TrajectoryFeatures,
        data.frame(
          Sujeto = s,
          Ensayo = e,

          Choice = unique(
            data_se$ChosenCondition
          ),

          FirstAOI = net$FirstAOI,
          FirstTransition = net$FirstTransition,
          First3AOIs = net$First3AOIs,

          Nodes = gorder(
            net$graph
          ),

          Edges = gsize(
            net$graph
          ),

          Entropy = net$Entropy,

          stringsAsFactors = FALSE
        )
      )

    }
  }
}

summary(
  TrajectoryFeatures
)
```

# Appendix B. AOI visitation frequencies

```{r aoi-frequency-table}
all_aois <- bind_rows(
  lapply(
    NetworkList,
    function(x){

      data.frame(
        AOI = names(
          x$node_freq
        ),
        Visits = as.numeric(
          x$node_freq
        )
      )

    }
  )
)

freq_tab <- all_aois %>%
  group_by(AOI) %>%
  summarise(
    Visits = sum(Visits)
  ) %>%
  arrange(
    desc(Visits)
  )

kable(
  freq_tab,
  caption = "AOI visitation frequencies."
)
```

```{r aoi-frequency-figure}
ggplot(
  freq_tab,
  aes(
    reorder(AOI, Visits),
    Visits
  )
)+
  geom_col(
    fill = "darkgreen"
  )+
  coord_flip()+
  theme_minimal()+
  labs(
    x = "AOI",
    y = "Visits"
  )
```

# Appendix C. AOI-to-AOI transition heatmap

## Aggregate transitions

```{r aggregate-transitions}
all_edges <- bind_rows(
  lapply(
    NetworkList,
    function(x){
      x$edges
    }
  )
)

agg_edges <- all_edges %>%
  count(
    from,
    to,
    name = "weight"
  )
```

## Transition matrix

```{r transition-matrix}
mat <- xtabs(
  weight ~ from + to,
  data = agg_edges
)

mat
```

## Hierarchical clustering heatmap

```{r heatmap}
pheatmap(
  mat,
  clustering_distance_rows = "euclidean",
  clustering_distance_cols = "euclidean",
  clustering_method = "ward.D2",
  color = colorRampPalette(
    c(
      "white",
      "darkgreen"
    )
  )(100)
)
```

## Export figure

```{r export-heatmap, eval=FALSE}
png(
  "F2.png",
  width = 5000,
  height = 3500,
  units = "px",
  res = 600
)

pheatmap(
  mat,
  clustering_distance_rows = "euclidean",
  clustering_distance_cols = "euclidean",
  clustering_method = "ward.D2",
  color = colorRampPalette(
    c(
      "white",
      "darkgreen"
    )
  )(100)
)

dev.off()
```

# Appendix D. Multiverse analysis

## Chi-square tests

```{r chisquare-tests}
tabla_firstAOI <- table(
  TrajectoryFeatures$FirstAOI,
  TrajectoryFeatures$Choice
)

test_firstAOI <- chisq.test(
  tabla_firstAOI
)

tabla_firstTransition <- table(
  TrajectoryFeatures$FirstTransition,
  TrajectoryFeatures$Choice
)

test_firstTransition <- chisq.test(
  tabla_firstTransition
)

tabla_first3 <- table(
  TrajectoryFeatures$First3AOIs,
  TrajectoryFeatures$Choice
)

test_first3 <- chisq.test(
  tabla_first3
)
```

## ANOVAs

```{r anovas}
anova_nodes <- summary(
  aov(
    Nodes ~ Choice,
    data = TrajectoryFeatures
  )
)

anova_edges <- summary(
  aov(
    Edges ~ Choice,
    data = TrajectoryFeatures
  )
)

anova_entropy <- summary(
  aov(
    Entropy ~ Choice,
    data = TrajectoryFeatures
  )
)
```

## Summary table

```{r multiverse-table}
multiverse_results <- data.frame(
  Universe = c(
    "FirstAOI",
    "FirstTransition",
    "First3AOIs",
    "Nodes",
    "Edges",
    "Entropy"
  ),

  Test = c(
    "Chi-square",
    "Chi-square",
    "Chi-square",
    "ANOVA",
    "ANOVA",
    "ANOVA"
  ),

  Pvalue = c(
    test_firstAOI$p.value,
    test_firstTransition$p.value,
    test_first3$p.value,
    anova_nodes[[1]][
      "Choice",
      "Pr(>F)"
    ],
    anova_edges[[1]][
      "Choice",
      "Pr(>F)"
    ],
    anova_entropy[[1]][
      "Choice",
      "Pr(>F)"
    ]
  )
)

kable(
  multiverse_results,
  digits = 4,
  caption = "Multiverse analysis of visual exploration."
)
```

# Appendix E. Multinomial predictive models

```{r prepare-choice}
TrajectoryFeatures$Choice <- as.factor(
  TrajectoryFeatures$Choice
)
```

## Model estimation

```{r multinomial-models}
m1 <- multinom(
  Choice ~ Entropy,
  data = TrajectoryFeatures,
  trace = FALSE
)

m2 <- multinom(
  Choice ~ Entropy + FirstAOI,
  data = TrajectoryFeatures,
  trace = FALSE
)

m3 <- multinom(
  Choice ~ Entropy + FirstTransition,
  data = TrajectoryFeatures,
  trace = FALSE
)

m4 <- multinom(
  Choice ~ Nodes + Edges + Entropy,
  data = TrajectoryFeatures,
  trace = FALSE
)
```

## Comparative performance table

```{r multinomial-results}
results_models <- data.frame(
  Model = c(
    "M1",
    "M2",
    "M3",
    "M4"
  ),

  Predictors = c(
    "Entropy",
    "FirstAOI + Entropy",
    "FirstTransition + Entropy",
    "Nodes + Edges + Entropy"
  ),

  McFadden_R2 = c(
    pR2(m1)["McFadden"],
    pR2(m2)["McFadden"],
    pR2(m3)["McFadden"],
    pR2(m4)["McFadden"]
  ),

  AIC = c(
    AIC(m1),
    AIC(m2),
    AIC(m3),
    AIC(m4)
  )
)

results_models <- results_models %>%
  mutate(
    McFadden_R2 = round(
      McFadden_R2,
      4
    ),
    AIC = round(
      AIC,
      1
    )
  )

kable(
  results_models,
  caption =
    "Comparative performance of multinomial models predicting booking decisions from alternative representations of visual exploration."
)
```